The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-30 paper.tex
Added numbers to SonRise, and added more detailed comparison of SonAngle vs SonRise

\ex. Desiderata for a sonority contour similarity metric \label{desiderata}
     \a. The more positive the gradient of the sonority contour from C\textsubscript{1} to C\textsubscript{2}, the smaller the faithfulness cost.
     \b. For a given sonority distance between C\textsubscript{1} and C\textsubscript{2}, the more sonorous C\textsubscript{1} and C\textsubscript{2} are, the smaller the faithfulness cost.

@@

\vspace{-3em}
\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityanglehierarchycompressed}}

@@

\ex.    \begin{center}
    \begin{tabular}{c c | c c | c}
    C\textsubscript{1}C\textsubscript{2}   & Rise & C\textsubscript{1}V & Rise & \textsc{SonRise} Distance\\ \hline
    NT &  -2  &  N\textipa{1}T & 3 & $1-\frac{1-3}{6-3}=1.7$ \\
    TT &  0   &  T\textipa{1}T & 5 & $1-\frac{1-1}{6-1}=1.0$ \\
    TN &  2   &  T{\textipa{1}}N & 5 & $1-\frac{3-1}{6-1}=0.6$ \\ 
    \end{tabular}
    \end{center}

@@

\vspace{-3em}
\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityrisehierarchycompressed}}

Observe that of the two desiderata for such a metric, {\sc Sonority Rise} fulfills the first but not the second. Consider the falling sonority clusters. RN, NF and FT, which all have a sonority distance of -1, have a smaller {\sc Sonority Rise} than NT and RF, with a sonority distance of 2, followed by RT with a sonority distance of 3.

However, of the three, RN has the largest {\sc Sonority Rise}, and therefore is expected to be more resistant to epenthesis than NF and FT. If RN undergoes epenthesis, therefore, we predict FT and NF to do the same.

@@

@@

I will formalise this restriction with the constraint {\sc Dep}(+sonorant). The epenthesis of a vowel between two obstruents, which bear the feature [$-$sonorant], necessitates the insertion of a new [+sonorant] feature, whereas when a vowel is epenthesised adjacent to a sonorant, the sonorant's [+sonorant] feature spreads and is shared between the vowel and the sonorant.

>>>>

\ex. Desiderata for a sonority contour faithfulness metric \label{desiderata}
     \a. The more positive the gradient of the sonority contour from C\textsubscript{1} to C\textsubscript{2}, the smaller the faithfulness cost. \label{desideratum_a}
     \b. For a given sonority distance between C\textsubscript{1} and C\textsubscript{2}, the more sonorous C\textsubscript{1} and C\textsubscript{2} are, the smaller the faithfulness cost. \label{desideratum_b}

@@

\vspace{-3em}
\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityanglehierarchycompressednumbers}}

@@

\ex. Sample calculations of {\sc SonRise}
   \begin{center}
    \begin{tabular}{c c | c c | c}
    C\textsubscript{1}C\textsubscript{2}   & Rise & C\textsubscript{1}V & Rise & \textsc{SonRise} Distance\\ \hline
    NT &  -2  &  N\textipa{1}T & 3 & $1-\frac{1-3}{6-3}=1.7$ \\
    TT &  0   &  T\textipa{1}T & 5 & $1-\frac{1-1}{6-1}=1.0$ \\
    TN &  2   &  T{\textipa{1}}N & 5 & $1-\frac{3-1}{6-1}=0.6$ \\ 
    \end{tabular}
    \end{center}

@@

\vspace{-3em}
\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityrisehierarchycompressednumbers}}

Observe that of the two desiderata for sonority contour metrics outlined in \ref{desiderata}, {\sc Sonority Rise} fulfills the first but not the second. Holding C\textsubscript{1} constant, closing the sonority gap with C\textsubscript{2} does result in a smaller {\sc SonRise} --- compare RT, RF and RN, for example.

However, out of the three clusters with a sonority distance of -1, RN has the largest {\sc Sonority Rise}, and therefore is expected to be more resistant to epenthesis than NF and FT. This is contrary to criterion \ref{desideratum_b}. If RN undergoes epenthesis, therefore, we predict FT and NF to do the same.

@@

Let us do a more detailed comparison of the two cluster hierarchies:

\ex. Detailed comparison of {\sc Sonority Rise} vs {\sc Sonority Angle} hierarchies:

\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityanglehierarchycompressed}}

\vspace{-2em}
\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityrisehierarchycompressed}}

Within the rising sonority clusters, the predictions of the two hierarchies are identical.
The differences in predictions are as follows:

\ex. \a. More sonorous plateaus are more likely to undergo epenthesis under the {\sc Sonority Angle} metric, whereas all the level clusters should have equal likelihood of being broken up by epenthesis.
     \b. RN is more likely to undergo epenthesis than NF under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RN is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. NF is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than RT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than NT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RT and NT are equally likely to undergo epenthesis under {\sc Sonority Angle}, but NT is more likely to undergo epenthesis than RF under {\sc Sonority Rise}.

@@

I will formalise this restriction with the constraint {\sc Dep}(+sonorant).

\ex. {\sc Dep}(+sonorant): do not insert a sonorant region between two obstruents.

The epenthesis of a vowel between two obstruents, which bear the feature [$-$sonorant], necessitates the insertion of a new [+sonorant] feature, whereas when a vowel is epenthesised adjacent to a sonorant, the sonorant's [+sonorant] feature spreads and is shared between the vowel and the sonorant.
